IPv6 over Low-Power Wireless Personal Area Networks\\ MSc. Thesis Proposal

### 2010-01-17 doc/msc-proposal/msc-routing.tex
proposal: corrected ieee 802.15.4

The survey uses five criteria derived from a set of requirements for routing in low power and lossy networks:  routing state, loss response, control cost, link cost, and node cost. The routing state criterion indicates whether routing state scales reasonably within the memory resources of low-power nodes. Routing state that scales linearly with the size of the network or a node's neighbourhood fails and passes if scales with the number of destinations. The loss response indicates whether the protocol localizes responses to link failures with no triggering of global network re-optimization. Protocols which require many link changes to propagate across the entire network fail. The control cost criterion defines constraints on control traffic, in order to discover a topology. The link and node cost specify how a protocol chooses routes for data packets to take through the network. A protocol passes these criteria if it provides a mechanism allowing link and node properties to be considered when choosing routes.

Table \ref{table:routing.prot.survey} summarizes the survey showing which of existing protocols meet the criteria described above. The detailed analysis is given in \cite{draft-protocols-07}. For each of these criteria, the value "pass" indicates that a protocol has satisfactory performance.  The value "fail" corresponds to not acceptable performance, which means that the protocol does not meet the criterion. Finally, a "?" means that a protocol would require a supplementary document specifying how a protocol should behave. As can be seen from the table, no existing IETF protocol meets the described criteria. Therefore, the survey concludes that a new protocol specification documents have to be defined for a LLN routing protocol.

@@

\section{Routing Protocol for Low power and Lossy networks}\label{sec:rout.rpl}

\subsection{Routing Metrics}
Historically, IGP such as OSPF ([RFC2328]) and IS-IS ([RFC1195]) have used quantitative static link metric

>>>>

The survey uses five criteria derived from a set of requirements for routing in low power and lossy networks:  routing state, loss response, control cost, link cost, and node cost. The routing state criterion indicates whether routing state scales reasonably within the memory resources of low-power nodes. Routing state that scales linearly with the size of the network or a node's neighbourhood fails, and passes if scales with the number of destinations. The loss response indicates whether the protocol localizes responses to link failures with no triggering of global network re-optimization. Protocols which require many link changes to propagate across the entire network fail. The control cost criterion defines constraints on control traffic required to discover a topology. The link and node cost specify how a protocol chooses routes for data packets to take through the network. A protocol passes these two criteria if it provides a mechanism allowing both link and node properties to be considered when choosing routes.

Table \ref{table:routing.prot.survey} summarizes the survey showing which of existing protocols meet the criteria described above. The detailed analysis is given in the document \cite{draft-protocols-07}. For each of these criteria, the value "pass" indicates that a protocol has satisfactory performance.  The value "fail" corresponds to not acceptable performance, which means that the protocol does not meet the criterion. Finally, a "?" means that a protocol would require a supplementary document specifying how a protocol should behave. As can be seen from the table, no existing IETF protocol meets all described criteria. Therefore, the survey concludes that new protocol specification documents have to be defined for a LLN routing protocol.

@@

\section{Overview of IPv6 Routing Protocol for LLNs}\label{sec:rout.rpl}
The IPv6 Routing Protocol for Low power and Lossy Networks (RPL) introduces a notion of a Directed Acyclic Graph (DAG), which is a directed graph with no cycles exist. All edges of a DAG are oriented toward and terminating at one or more root nodes uniquely identified by the DAGID. A Destination Oriented DAG (DODAG) is a DAG rooted at a single destination. While constructing a DODAG various routing metrics and optimization objectives can be used to compute the shortest paths. The set of supported link and node metrics used in the RPL is specified in the internet draft \cite{draft-routing-metrics-04}.

A DAG instance is a set of possibly multiple DODAGs. The routing metrics and objectives for a certain DAG instance are specified by an Objective Function (OF). Each DAG instance constructs a routing topology optimized for its OF. Traffic that belongs to a specific DODAG instance is marked in the flow label of the IPv6 header. A network may have more than one DAG instance which operate independently. This allows to applications to tag traffic to follow an appropriate DAG instance, i.e., optimized for low latency or low energy.

\subsection{Messages}
The protocol defines the RPL Control Message, which is a new ICMPv6 message. The format of this message is shown in Figure \ref{fig:rpl.control.message}.
\begin{figure}[htp]
\begin{mylisting}
\begin{verbatim}
   0                   1                   2                   3
   0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1
   +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
   |     Type      |     Code      |          Checksum             |
   +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
   |                                                               |
   +                         Message Body                          +
   |                                                               |
\end{verbatim}
\end{mylisting}
\caption{RPL Control Message}\label{fig:rpl.control.message}
\end{figure}

The Type for the RPL Control message has the value of 155. The Code identifies the RPL Control Messages as follows:  0x01 -- DAG Information Solicitation; 0x02: DAG Information Object; 0x04: Destination Advertisement Object.

DAG Information Object (DIO) message exchanges are used by RPL nodes to construct and maintain DODAGs. A DIO message identifies the DAG instance, the DAGID, and the values used to compute the DAG instance's objective function. The DIO also includes a measure derived from the location of the node within the DODAG, the rank, used to determine the position of the node relative to other nodes and avoid loops. The exact calculation of the rank is left to the Objective Function and  depends on the parents of the node. The DAG Information Solicitation (DIS) message is used to solicit a DAG Information Object from a RPL node.

DIO messages allow to establish upward routes, i.e., the routes in the direction from leaf nodes towards DODAG roots.  In order to establish downward routes Destination Advertisement Object (DAO) messages are used. The DAO propagates destination information upwards along the DODAG.

The detailed format of the messages is specified in the internet draft \cite{draft-rpl-04}.

\subsection{Basic Operation}
A node's neighbor set is the nodes that it can reach with a link-local multicast. A node that is not a DODAG root may maintain multiple DAG parents for a single DAG instance. DAG parents of a node is a subset of the set of       its neighbors. A node may implement different policies for selecting its DAG parents.

DAG discovery allows a node to join a DODAG by discovering neighbors that are members of the DODAG.  Some nodes are preconfigured to be DODAG roots. Each node maintains a timer that governs when to advertise node presence by sending link-local multicast DIO messages. The rate at which DIO messages are sent varies in response to stability or detection of routing inconsistencies. Nodes listen for DIOs and use their information to join a new DODAG, or to maintain an existing DODAG, according to the specified Objective Function. DODAG discovery avoids loops by constraining how and when nodes can increase their rank.

The destination advertisement mechanism supports the dissemination of routing state required to support traffic flows down along the DODAG, from the DODAG root toward nodes. Destinations disseminated with the destination advertisement mechanism may be prefixes, individual hosts, or multicast listeners.  The mechanism supports nodes with varying capabilities. Nodes that are capable to maintain routing state may inspect destination advertisements and learn hop-by-hop routing state toward destinations populating their routing tables with the learned routes. They may also learn necessary piecewise source routes to traverse regions of the LLN that do not maintain routing state. Route aggregation on known destinations can be accomplished by such nodes before emitting Destination Advertisements. The nodes that are incapable of storing routing state forward destination advertisements, attaching a next-hop address to the reverse route used to support the construction of piecewise source routes.

A comprehensive description of the RPL is given in the internet draft \cite{draft-rpl-04}
